---
name: "Priyanka Srinivasan"
title: "p8105_hw1_ps3070"
output: github_document
---

**Problem 1**

First, I will load the tidyverse library.
```{r adding tidyverse library, results = "hide"}
library(tidyverse)
```

Next, I will create a dataframe comprised of:
* a random sample of size 8 from a standard Normal distribution
* a logical vector indicating whether elements of the sample are greater than 0
* a character vector of length 8
* a factor vector of length 8, with 3 different factor “levels”

First, I set the seed so my results would be reproducible.
```{r set seed}
set.seed(123)
```


```{r create dataframe, results="hide"}
prob1_df = tibble(
  random_sample = rnorm(8),
  vec_logical = random_sample > 0, #This creates a logical vector the same length as random_sample with the value True when this condition is satisfied and False when it isn't.
  vec_character = c("abc", "def", "ghi", "jkl", "mno", "pqr", "stu", "vwx"),
  vec_factor = factor(c("p","r","i","p","r","i","p","r"))
)
```

The following code chunk attempts to take the mean of each vector.
```{r take mean}
mean(pull(prob1_df, random_sample))
mean(pull(prob1_df, vec_logical))
mean(pull(prob1_df, vec_character))
mean(pull(prob1_df, vec_factor))
```

The mean of `random_sample` and `vec_logical` are possible to compute because they are numeric and logical variables. The logical variable values will be interpreted as `True` = 1 and `False` = 0, so the mean will be an average of these results. The means for vectors `vec_character` and `vec_factor` cannot be computed because, to calculate the mean, R requires the argument to be numeric or logical.



The following code chunk will use `as.numeric()` to try to convert factor, logical, and character variables to numeric.

```{r numeric conversion, results="hide"}
fac_to_num = as.numeric(pull(prob1_df, vec_factor))
char_to_num = as.numeric(pull(prob1_df, vec_character))
log_to_num = as.numeric(pull(prob1_df, vec_logical))
```

Factor variables will convert to numeric variables based on the integer values that R assigns to them. Every time a factor vector appears, R will assign the same number to it. Since I created the factor vector with letters, the integers are assigned starting with 1, in alphabetical order. 

The logical variable values will be assigned as `True` = 1 and `False` = 0.

R tries to coerce variables of a different type (i.e. not numeric or logical) which resultst in NAs. Since the character vector is not based on numbers/integers, we are unable to convert them to numeric. Because of that, I will try to convert the character variable into a factor variable, and then into a numeric variable:

```{r convert character to numeric, results="hide"}
char_to_num2 = as.numeric(as.factor(pull(prob1_df, vec_character)))
```
Now, the warning message is not received and this produces a string of numbers. Conversion of a character variable to factor and then to numeric was successful.


In the next code chunk, I will:

* convert the logical vector to numeric, and multiply the random sample by the result
* convert the logical vector to a factor, and multiply the random sample by the result
* convert the logical vector to a factor and then convert the result to numeric, and multiply the random sample by the result

```{r conversions, results="hide"}
as.numeric(pull(prob1_df, vec_logical)) * pull(prob1_df, random_sample)
as.factor(pull(prob1_df, vec_logical)) * pull(prob1_df, random_sample)
as.numeric(as.factor(pull(prob1_df, vec_logical))) * pull(prob1_df, random_sample)
```

The first and third operations are successful (logical ot numeric, and logical to factor and tthen numeric. However, when converting the logical vector to a factor variable and then multiplying random sample (numeric) by the result, the console displays a message saying "not meaningful for factors" and displays a string of NAs. Only after converting the factor variable to numeric will R multiply this with a numeric variable.



**Problem 2**

I created a data frame comprised of:

* x: a random sample of size 500 from a standard Normal distribution
* y: a random sample of size 500 from a standard Normal distribution
* A logical vector indicating whether x + y > 1
* A numeric vector created by coercing the above logical vector
* A factor vector created by coercing the above logical vector

```{r setting seed}
set.seed(123)
```

```{r create dataframe 2, results="hide"}
prob2_df = tibble(
  x = rnorm(500),
  y = rnorm(500),
  vec_log = (x + y) > 1,
  vec_num = as.numeric(vec_log),
  vec_fac = as.factor(vec_log)
)
```


The dataset includes `r nrow(prob2_df)` rows and `r ncol(prob2_df)` columns, indicating that there are 5 vectors with 500 values. The mean of x is `r mean(pull(prob2_df, x))`. The median of x is `r median(pull(prob2_df, x))`. The standard deviation of x is `r sd(pull(prob2_df, x))`. The proportion of cases for which (x + y) > 1 is the same as the proportion of cases for which the logical vector is 'True', which is `r length(which(pull(prob2_df, vec_log)))/length(pull(prob2_df, vec_log))`.



Now, I will make a scatterplot of `y` vs `x` using the logical variable `vec_log`, equaling `(x + y) > 1`, as well as the numeric and factor variables created from `vec_log`.

The logical vector `vec_log` and the factor vector `vec_fac` are based on binary values. For the logical vector, values are `True` and `False`, and for the factor vector, values are `0` and `1`. For the numeric vector even though the data values only include `0` and `1`, R will still assume it is continuous. The legends for each plot are created accordingly - for `vec_log` and `vec_fac`, the legends show `FALSE` and `TRUE`, but for `vec_num`, there is a gradient from `0` to `1`, though the plot itself shows only two colors. Using `geom_point`, I decreased the size of the points tot `0.5` so that less points overlapped.

```{r create scatterplots}
plot_log <- ggplot(prob2_df, aes(x = x, y = y, color = vec_log)) + geom_point(size = 0.5)
plot_num <- ggplot(prob2_df, aes(x = x, y = y, color = vec_num)) + geom_point(size = 0.5)
plot_fac <- ggplot(prob2_df, aes(x = x, y = y, color = vec_fac)) + geom_point(size = 0.5)

plot_log
plot_num
plot_fac
```

```{r export first scatterplot}
ggsave("plot_log.pdf", height = 4, width = 6)
```
